taxadb: A High-Performance Local Taxonomic Database Interface

### 2020-02-12 paper/manuscript.Rmd
final spell check

```{r echo = FALSE, cache = FALSE}
joined %>%
  printtable(caption = "Example IUCN and trait data joined directly on scientific name showing only one match. While common, joining on scientific name does not account for taxnomic inconsistiencies between databases and therefore results in seemingly very little overlap in species representation between the two.")
```

@@

```{r cache = FALSE, echo = FALSE}
## Just for pretty-printing
joined %>%  
  tidyr::replace_na(list(category = "-", elton_name = "-", iucn_name = "-")) %>%
  select(elton_name, iucn_name, mass, category, id) %>%
  printtable(caption = "Example IUCN and trait data joined on taxonomic ID. Multilple species have a different scientific name in the Elton and IUCN Redlist databases but can be match based on their COL taxonomic ID.")
```

@@

The full taxonomic record in the database can also be directly accessed by `taxa_tbl()`, allowing for whole-database queries that are not possible through the API or web interface of many providers.  For example, we can easily check the coverage of accepted species names in each of the classes of vertabrates within the Catalogue of Life (Table 7):

@@

```{r echo = FALSE}
 verts %>%
  printtable(caption = "taxadb also provides direct access to the database, allowing dplyr or SQL queries which can compute across the entire dataset, such as counting accepted species in all vertabrate classes shown here.  This kind of query is effectively impossible in most REST API-based interfaces.")
```

@@

Lastly, we note that local database design used in `taxadb` is not unique to taxonomic names.  Despite the rapid expansion of REST API-based interfaces to ecological data [@ropensci], in our experience, much of the data relevant to ecologists and evolutionary biologists today would be also be ameniable to the local database design.  The local database approach is much easier for data providers (who can leverage static scientific database repositories instead of maintaining REST servers) and often much faster for data consumers.

>>>>

```{r echo = FALSE, cache = FALSE}
joined %>%
  printtable(caption = "Example IUCN and trait data joined directly on scientific name showing only one match. While common, joining on scientific name does not account for taxonomic inconsistencies between databases and therefore results in seemingly very little overlap in species representation between the two.")
```

@@

```{r cache = FALSE, echo = FALSE}
## Just for pretty-printing
joined %>%  
  tidyr::replace_na(list(category = "-", elton_name = "-", iucn_name = "-")) %>%
  select(elton_name, iucn_name, mass, category, id) %>%
  printtable(caption = "Example IUCN and trait data joined on taxonomic ID. Multiple species have a different scientific name in the Elton and IUCN Redlist databases but can be match based on their COL taxonomic ID.")
```

@@

The full taxonomic record in the database can also be directly accessed by `taxa_tbl()`, allowing for whole-database queries that are not possible through the API or web interface of many providers.  For example, we can easily check the coverage of accepted species names in each of the classes of vertebrates within the Catalogue of Life (Table 7):

@@

```{r echo = FALSE}
 verts %>%
  printtable(caption = "taxadb also provides direct access to the database, allowing dplyr or SQL queries which can compute across the entire dataset, such as counting accepted species in all vertebrate classes shown here.  This kind of query is effectively impossible in most REST API-based interfaces.")
```

@@

Lastly, we note that local database design used in `taxadb` is not unique to taxonomic names.  Despite the rapid expansion of REST API-based interfaces to ecological data [@ropensci], in our experience, much of the data relevant to ecologists and evolutionary biologists today would be also be amenable to the local database design.  The local database approach is much easier for data providers (who can leverage static scientific database repositories instead of maintaining REST servers) and often much faster for data consumers.
